Reproducible Research: Peer assignment 1
========================================================

## Dependencies
```{r}
library(ggplot2)
library(plyr)
library(timeDate)
library(lattice)
```

## Loading and preprocessing the data

Read in the csv file and convert dates to to the right format.
```{r}
data <- read.csv("activity.csv")
data$date <- as.Date(as.character(data$date, format="%Y/%m/%d"))
```

## What is mean total number of steps taken per day?

1. Histogram of the total number of steps taken each day
```{r}
qplot(data$date, data=data, weight=data$steps, geom="histogram", binwidth = 1) + 
  xlab("Date") + 
  ylab("Number of steps") + 
  ggtitle("Total number of steps per day")
```

2. Transform data using plyr:
```{r}
data.sum.steps.per.day <- ddply(data, c("date"), function(df)c(sum(df$steps, na.rm=TRUE)))
```

The mean total number of steps taken per day (ignoring missing values) is :
```{r}
mean(data.sum.steps.per.day$V1, na.rm=TRUE)
```
The median total number of steps taken per day (ingoring missing values) is : 
```{r}
median(data.sum.steps.per.day$V1, na.rm=TRUE)
```

## What is the average daily activity pattern?

1. Transform data using plyr:
```{r}
num_days <- length(unique(data$date))
data.avg.steps.per.interval <- ddply(data, c("interval"), function(df)c(sum(df$steps, na.rm=TRUE)/num_days))
```

Plot the time series:
```{r}
qplot(data.avg.steps.per.interval$interval, data.avg.steps.per.interval$V1, geom="line") + 
  xlab("5-minute interval") + 
  ylab("Number of steps averaged across all days") + 
  ggtitle("Average daily activity pattern")
```

2. The 5 minute interval that contains the maximum number of steps (averaged across all days is):
```{r}
data.avg.steps.per.interval[which.max(data.avg.steps.per.interval$V1), c("interval")]
```

## Imputing missing values

1. The total number of missing values in the dataset is:
```{r}
length(data[is.na(data$steps), c("steps")])
```

2-3. Impute missing values
Copy the data and for convert each NA to the average number of steps for that interval:
```{r}
data.impute <- data
na.index <- which(is.na(data.impute$steps))
for(index in na.index ){
  data.impute$steps[index] <- data.avg.steps.per.interval[data.avg.steps.per.interval$interval==data.impute$interval[index] , c("V1")]
}
```


4. Histogram of the total number of steps taken each day
```{r}
qplot(data.impute$date, data=data.impute, weight=data.impute$steps, geom="histogram", binwidth = 1) + 
  xlab("Date") + 
  ylab("Number of steps") + 
  ggtitle("Total number of steps per day")
```

Transform data using plyr:
```{r}
data.impute.sum.steps.per.day <- ddply(data.impute, c("date"), function(df)c(sum(df$steps)))
```

The mean total number of steps taken per day (ignoring missing values) is :
```{r}
mean(data.impute.sum.steps.per.day$V1)
```
The median total number of steps taken per day (ingoring missing values) is : 
```{r}
median(data.impute.sum.steps.per.day$V1)
```

Those estimates are not the same as those computed in the first part of the assignment.
As expected, the total number of steps has increased:
* Total number of steps before: `r sum(data$steps, na.rm=TRUE)`
* Total number of steps after: `r sum(data.impute$steps)`


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variables for weekdays/weekends and add it to the data frame with the imputed data:
```{r}
weekdays <- as.factor(isWeekday(data.impute$date))
levels(weekdays) <- c("weekend", "weekday")
data.impute <- cbind(data.impute, weekdays)

```

Transform the data into appropriate form:
```{r}
weekday.data <- data.impute[data.impute$weekdays=="weekday",]
weekend.data <- data.impute[data.impute$weekdays=="weekend",]
num_weekdays <- length(unique(weekday.data$date))
num_weekends <- length(unique(weekend.data$date))
avg.steps.per.interval.weekday <- ddply(weekday.data, c("interval"), function(df)c(sum(df$steps, na.rm=TRUE)/num_weekdays))
avg.steps.per.interval.weekend <- ddply(weekend.data, c("interval"), function(df)c(sum(df$steps, na.rm=TRUE)/num_weekends))
```

Plot:
```{r}
par(mfrow = c(2, 1), mar = c(3, 3, 1, 1))

  plot(avg.steps.per.interval.weekend$interval, avg.steps.per.interval.weekend$V1, type="n", xlab="Interval", ylab="Average number of steps across all days", main = "Weekend")
  lines(avg.steps.per.interval.weekend$interval, avg.steps.per.interval.weekend$V1, col="blue")

  plot(avg.steps.per.interval.weekday$interval, avg.steps.per.interval.weekday$V1, type="n", xlab="Interval", ylab="Average number of steps across all days", main = "Weekdays")
  lines(avg.steps.per.interval.weekday$interval, avg.steps.per.interval.weekday$V1, col="blue")
```
